---
title: "p8105_hw3_yl5214"
output: github_document
---

```{r}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

## Problem 2
```{r}
library(p8105.datasets)
data("brfss_smart2010")
view(brfss_smart2010)
```

```{r}
brfss=brfss_smart2010 |> 
  janitor::clean_names() |> 
  filter(topic=='Overall Health') |> 
  filter(response %in% c('Excellent', 'Very good','Good','Fair','Poor')) |> 
  mutate(
    response= factor(response, levels = c("Poor", "Fair", "Good", "Very good","Excellent"), ordered = TRUE)
  ) |> 
rename(state =locationabbr) |> 
rename(county =locationdesc)
```

## In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
brfss_2002=brfss |> 
  filter(year=='2002') |> 
  group_by(state) |> 
  summarize(n_obs = n_distinct(county)) |> 
  filter(n_obs>=7)

brfss_2010=brfss |> 
  filter(year=='2010') |> 
  group_by(state) |> 
  summarize(n_obs = n_distinct(county)) |> 
  filter(n_obs>=7)
```
In 2002, `r print(pull(brfss_2002, state))` 6 states were observed at 7 or more locations. 
In 2010, `r print(pull(brfss_2010, state))` 14 states were observed at 7 or more locations.


## Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
brfss_excellent=brfss |> 
  filter(response == 'Excellent') |> 
  group_by (year, state) |> 
  summarise(
    average_value= mean(data_value, na.rm = TRUE),
    .groups = 'drop')|> 
  ggplot(aes (x= year, y= average_value, color=state))+
  geom_line()
  
```


##Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss_two=brfss |> 
  filter(
    year %in% c('2010', '2006')
         ) |> 
  filter(state == 'NY') |> 
  group_by (year) |> 
  ggplot(aes (x= response, y= data_value))+
  geom_boxplot()+
  facet_grid(. ~ year)
  
```

## Problem 3

##Load, tidy, merge, and otherwise organize the data sets. Your final dataset should include all originally observed variables; exclude participants less than 21 years of age, and those with missing demographic data; and encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).
```{r}
covar=
  read_csv('data/nhanes_covar.csv', 
           skip =4,
           col_types=
             cols(
               'SEQN'=col_factor()
             ))|> 
  janitor::clean_names() |> 
  filter(age>=21) |> 
  drop_na(bmi) |> 
  drop_na(education) |> 
  mutate(
    sex= case_match(
      sex,
      1~ 'male',
      2 ~'female'
    )
  ) |> 
  mutate(
    education=case_match(
      education,
      1~'less than high school',
      2~'high school equivalent',
      3~'more than high school'
    )
  )
```

```{r}
accel=
  read_csv('data/nhanes_accel.csv', 
           col_types=
             cols(
               'SEQN'=col_factor()
             ))|> 
  janitor::clean_names()

merge=
  left_join(covar,accel, by = 'seqn')

```

## Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items.
```{r}
merge |>  
  count(education,sex) |> 
    pivot_wider(
    names_from =sex,
    values_from= n
)
knitr::kable
```

